What's in a Domain? Analysis of URL Features

### 2023-08-05 paper/paper.tex
Typos and formatting

Processing URLs is key component of many tasks that involve
analysing internet data. In many applications the URL is the central piece of 
information available because the demands of the task require immediate analysis.
In applications like malicious wesite detection the URL needs to be processed
in a rapid and efficient manner to provide utility\cite{Ma2009b}.
Common use cases for URL centered analysis include security analysis of potenital phishing 
attacks\cite{Garera2007,Basnet2012,Basnet2014,Mamun2016,Verma2017,Vazhayil2018,Tupsamudre2019,Sirigineedi2020,Li2020}, 
and identification of pages that host malware or viruses \cite{Canali2011,Mamun2016}.
There are also applications to online advertising,
including anticipation of conversions \cite{Qiu2020}, contextual analysis of
the content\cite{Kan2004,Shih2004,Baykan2009,Meshkizadeh2010,Hernandez2012,Arya2016}, relevance \cite{Kan2005} 
or language\cite{Baykan2013} of webpages. 
Content categorisation has also been applied to spam web pages using URLs alone for the sake
of search results filtering\cite{Chung2010,Hassan2008}.

@@

Beyond the domain, the URL consists of a sequence of words, categories, identfiers, dates
or domain specific abreviations. This sequence can indicate the psychological elements of how
information is categeoried, or the internal logic of an application that generates or presents
content dynamically.

@@

Some authors have 
emphasized that URLs are sequences of characters and require feature extraction methods that respect 
this sequential nature\cite{Le2018,Vazhayil2018}. They have naturally turned to developing neural network approaches
based on convolutional or recurrent layers for learning these sequential structures.
However, the ordering of key features within URLs is predominantly fixed as shown in Figure \ref{fig:url_structure}.
The core elements of URL structure occur at fixed positions within the sequence, in contrast to human language
sequences which are built from flexible grammars that allow variable positioning of most key elements. The
inherent structure of URLs has been demonstrated to be a source of information that can be exploited in 
algorithm design \cite{Shih2004}. Furthermore,
there are psychological and social elements to URL construction (such as the use of common brand names in subdomains
for phishing attack URLS \cite{Tupsamudre2019}), which emphasize the utility of feature engineering techniques that
respect the inherent structure of URLs.

@@

In this work we develop a feature engineering library that maintains an ontology of URL features
related to underlying structure of the URL. In addition we provide a method for visualising the importance of features
across the URL structure. These techniques can prove useful for analysing features for problems that
consist of moving targets and require broad sets of potential features and efficient feature selection \cite{Basnet2012}.
We then present an open source package for generating features within these
ontologocal classes and apply them to a variety of tasks. We present our results as a
cross-task evaluation of URL features. The strength of our approach lies in the ability to both understand how
machine learning models use URLs in decision making, this is opposed to the black-box problem with neural network
models that treat the URL as a text sequence. In addition, the cross-task evaluation we perform provides insight into
what works within different domains. This is critical as many machine learning problems that rely on URLS require
models built on relatively small datasets, so that new patterns of fraud and behaviour can be reacted to in a timely
fashion.

@@

The structural components of a URL (as illustrated in Figure \ref{fig:url_structure}) are listed in Table \ref{tab:components}.
We include an overview the information source in each of those components and highlight how these components relate to
functional machine learning features.

@@

We design a feature extraction library for URLs that generates features specific to each component of the
URL independently using a naming convention that permits sepearation and grouping for analysis.
Previous work has presented datasets for specific problems where the features
are provided with a similar structural grouping\cite{Vrban2020}. However, we provide our feature extraction
application designed so that any specific subset of these features can be extracted and analysed 
for a specific task.

>>>>

Processing URLs is a key component of many tasks that involve
analysing internet data. In many applications the URL is the central piece of 
information available because the demands of the task require immediate analysis.
In applications like malicious wesite detection the URL needs to be processed
in a rapid and efficient manner to provide utility\cite{Ma2009b}.
Common use cases for URL centered analysis include security analysis of potential 
phishing attacks\cite{Garera2007,Basnet2012,Basnet2014,Mamun2016,Verma2017,Vazhayil2018,Tupsamudre2019,Sirigineedi2020,Li2020}, 
and identification of pages that host malware or viruses \cite{Canali2011,Mamun2016}.
There are also applications to online advertising,
including anticipation of conversions \cite{Qiu2020}, contextual analysis of
the content\cite{Kan2004,Shih2004,Baykan2009,Meshkizadeh2010,Hernandez2012,Arya2016}, relevance \cite{Kan2005} 
or language\cite{Baykan2013} of webpages. 
Content categorisation has also been applied to spam web pages using URLs alone for the sake
of search results filtering\cite{Chung2010,Hassan2008}.

@@

Beyond the domain, the URL consists of a sequence of words, categories, identifiers, dates
or domain specific abreviations. This sequence can indicate the psychological elements of how
information is categeoried, or the internal logic of an application that generates or presents
content dynamically.

@@

Some authors have emphasized that URLs are sequences of characters and require 
feature extraction methods that respect 
this sequential nature\cite{Le2018,Vazhayil2018}. 
They have turned to developing neural network approaches
based on convolutional or recurrent layers for learning these sequential structures.
However, the ordering of key features within URLs is predominantly fixed as shown 
in Figure \ref{fig:url_structure}. The core elements of URL structure occur at 
fixed positions within the sequence, in contrast to human language
sequences which are built from flexible grammars that allow variable positioning of 
most key elements. The inherent structure of URLs has been demonstrated to be a source 
of information that can be exploited in algorithm design \cite{Shih2004}. Furthermore,
there are psychological and social elements to URL construction (such as the use of 
common brand names in subdomains for phishing attack URLS \cite{Tupsamudre2019}), 
which emphasize the utility of feature engineering techniques that
respect the inherent structure of URLs.

@@

In this work we develop a feature engineering library that maintains an ontology of 
URL features related to underlying structure of the URL. In addition we provide a 
method for visualising the importance of features across the URL structure. 
These techniques can prove useful for analysing features for problems that
consist of moving targets and require broad sets of potential features and 
efficient feature selection \cite{Basnet2012}.
We present an open source package for generating features within these
ontologocal classes and apply them to a variety of tasks. We present our results as a
cross-task evaluation of URL features. The strength of our approach lies in the 
ability to both understand how machine learning models use URLs in decision making, 
as opposed to the black-box problem with neural network
models that treat the URL as a text sequence. In addition, the cross-task evaluation 
we perform provides insight into what works within different domains. 
This is critical as many machine learning problems that rely on URLS require
models built on relatively small datasets, so that new patterns of fraud and 
behaviour can be detected rapidly.

@@

The structural components of a URL (as illustrated in Figure \ref{fig:url_structure}) 
are listed in Table \ref{tab:components}.
We include an overview of the information source in each of those components and 
highlight how these components relate to machine learning features.

@@

We design a feature extraction library for URLs that generates features specific 
to each component of the URL independently using a naming convention that permits 
separation and grouping for analysis.
Previous work has presented datasets for specific problems where the features
are provided with a similar structural grouping\cite{Vrban2020}. However, we provide 
our feature extraction application designed so that any specific subset of these 
features can be extracted and analysed for a specific task.
